Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-11-30 inst/paper/main.tex
moving around

@@

Richards et al.~\cite{ecoop11} identified a number of frequently
occurring \eval patterns which were detected using simple regular
expressions on the \eval strings. In the case of R we match the
patterns on the recorded AST.

@@

\subsubsection{Recording the current call}
R provides a \c{match.call} function which can return the current
call. This is heavily used by the statistical model-fitting functions
that records the current call that is then reevaluated in the
different environment. Another pattern is to match the current call,
adjust some arguments and pass it to another function. For example the
\c{write.csv} function changes the \c{sep} argument in the matched
call and reevaluates itself as \c{write.table}:
%
\begin{lstlisting}
    Call <- match.call(expand.dots = TRUE)
    Call$sep <- ","; Call[[1L]] <- as.name("write.table")
    eval.parent(Call)
\end{lstlisting}

\subsubsection{Sandboxing} The combination of \eval and first class
environments can be used to construct a sandbox for evaluating
side-effecting expressions. The newly constructed environment can use
the current environment for its lexical scope. This provides the
evaluated expression access to all the bindings in the current scope
but restricts the side-effect to the newly constructed throwaway
environment. However, this sandboxing is limited because the
expression can reflectively access any environment in the call stack
or use the super assignment operator(\c{<<-}) to write to parent
scope. The \c{local} function which is part of Core R is an example of
this usage pattern.

\subsubsection{Dynamic Code Loading}
The combination of \parse and \eval is used to parse text from string
or file and evaluate in a custom environment. This pattern is used by
the \source Core R function to load R code from a file in the current
workspace. This is quite useful for loading functions for interactive
data analysis.

\subsubsection{Metaprogramming}
The combination of \eval and \substitute is used to do metaprogramming
in R. \substitute can be used to synthesize ASTs by stitching together
user inputs. \substitute walks over its argument AST and replaces the
symbols with their bindings in the specified environment and returns
this new AST which can then be evaluated in custom environments using
\eval.

\subsubsection{Convenience}
Eval is used to make some programming tasks easier (and/or shorter).
For example, in \c{data.table}, it is used to set R options from a
character vector
  \begin{lstlisting}
opts = c("datatable.verbose"="FALSE", ...) for (i in names(opts))
eval(parse(text=paste0("options(",i,"=",opts[i],")")))

>>>>

\mypara{Record Current Call.}
R provides a \c{match.call} function which can return the current
call. This is heavily used by the statistical model-fitting functions
that records the current call that is then reevaluated in the
different environment. Another pattern is to match the current call,
adjust some arguments and pass it to another function. For example the
\c{write.csv} function changes the \c{sep} argument in the matched
call and reevaluates itself as \c{write.table}:
%
\begin{lstlisting}
    Call <- match.call(expand.dots = TRUE)
    Call$sep <- ","; Call[[1L]] <- as.name("write.table")
    eval.parent(Call)
\end{lstlisting}

\mypara{Sandbox.} The combination of \eval and first class
environments can be used to construct a sandbox for evaluating
side-effecting expressions. The newly constructed environment can use
the current environment for its lexical scope. This provides the
evaluated expression access to all the bindings in the current scope
but restricts the side-effect to the newly constructed throwaway
environment. However, this sandboxing is limited because the
expression can reflectively access any environment in the call stack
or use the super assignment operator(\c{<<-}) to write to parent
scope. The \c{local} function which is part of Core R is an example of
this usage pattern.

\mypara{Code Loading.}
The combination of \parse and \eval is used to parse text from string
or file and evaluate in a custom environment. This pattern is used by
the \source Core R function to load R code from a file in the current
workspace. This is quite useful for loading functions for interactive
data analysis.

\mypara{Metaprogramming.}
The combination of \eval and \substitute is used to do metaprogramming
in R. \substitute can be used to synthesize ASTs by stitching together
user inputs. \substitute walks over its argument AST and replaces the
symbols with their bindings in the specified environment and returns
this new AST which can then be evaluated in custom environments using
\eval.

\mypara{Convenience.}
Eval is used to make some programming tasks easier (and/or shorter).
For example, in \c{data.table}, it is used to set R options from a
character vector
  \begin{lstlisting}
opts = c("datatable.verbose"="FALSE", ...) for (i in names(opts))
eval(parse(text=paste0("options(",i,"=",opts[i],")")))

@@

@@
